---
title: "College Acceptance Rate"
author: "Dhruv Raghunath"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
---


```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```


```{r load-libraries, include=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(lubridate)
library(here)
library(skimr)
library(janitor)
library(httr)
library(readxl)
library(vroom)
library(infer)
library(grid)
library(ggpubr)
library(rvest)
library(ggtext)
```

I will be performing a cluster analysis and K-means analysis on college acceptance data based on instate tuition rates and on `AcceptanceRate`, `Outofstate_tuition`, and `Graduation_rate`. This analysis was originally submitted as an assignment for my Business Analytics & Modeling class and has been slightly altered.

```{r}
colleges <- read.csv('Colleges.csv', header = T, as.is = T)
```

```{r}
#Creating an empty vector
colleges$AcceptanceRate <- c()

#Calculating acceptance rates for every college
for (i in 1:nrow(colleges)){
  colleges$AcceptanceRate[i] <- colleges$App_accepted[i]/colleges$App_received[i]
}

#Creating a scatter plot of instate tuition and acceptance rate
plot(colleges[, c('Instate_tuition', 'AcceptanceRate')]
     , xlab = 'Instate_tuition'
     , ylab = 'AcceptanceRate')

#Standardizing columns in the dataset
colleges_norm <- colleges
for (i in 2:ncol(colleges)){
  colleges_norm[,i] <- (colleges[, i] - mean(colleges[, i]))/sd(colleges[, i])
}

#Selecting only instate tuition and acceptance rate in the dataset
colleges_norm <- colleges_norm[,c(10,17)]

#Craeting a vector from 1 to 15
k_vec <- 1:15

#performing cluster analysis
avgdis <- c()
for (i in 1:length(k_vec)){
  km<- kmeans(colleges_norm, centers = k_vec[i])
  avgdis[i] <- mean(km$withinss)
}
avgdis

#Plotting the cluster analysis
plot(k_vec, avgdis,
     xlab="Number of clusters, k",
     ylab="Average within-clusters sum of squares")
```
Based on the plot above, I would use 5 clusters as the drop in sum of squares after the 5th cluster does is small



# performing a k-means analysis
```{r}
#Setting seed
set.seed(0)

km_subset<- kmeans(colleges_norm[,1:2], centers = 5)
```
The total number of observations in each cluster are: 59, 25, 135, 75, and 95.
The within cluster sum of squares by lusters are: 40.96, 20.79, 41.84, 24.61, and 38.57.
The cluster with 135 observations has the most homogeneity.


# Performing a cluster analysis using `AcceptanceRate`, `Outofstate_tuition`, and `Graduation_rate`
```{r}
colleges_norm2 <- colleges
for (i in 2:ncol(colleges)){
  colleges_norm2[,i] <- (colleges[, i] - mean(colleges[, i]))/sd(colleges[, i])
}

colleges_norm2 <- colleges_norm2[,c(11,16,17)]

avgdis2 <- c()
for (i in 1:length(k_vec)){
  km2<- kmeans(colleges_norm2, centers = k_vec[i])
  avgdis2[i] <- mean(km2$withinss)
}
avgdis2
plot(k_vec, avgdis2,
     xlab="Number of clusters, k",
     ylab="Average within-clusters sum of squares")

avgdis2[7]-avgdis2[8]
avgdis2[8]-avgdis2[9]
```
I would choose to create eight clusters. This is because the fall in average within cluster sum of squares from eight to nine clusters is only 6.28 while the fall from seven to eight clusters was nearly twice as much at 11.02.


# Citations
Raghunath, Dhruv. “DhruvRaghunath_A8.” 19 Nov. 2020. 

